---
title: "Congressional Districs of Wisconsin"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(maps)
library(mdsr)
library(fec16)
library(sf)
```


```{r, include=FALSE}
print(results_house, width = Inf)
results_house |>
  group_by(state, district_id) |>
  summarize(N = n())
```

```{r, include=FALSE}
district_elections <- results_house |>
  mutate(district = parse_number(district_id)) |>
  group_by(state, district) |>
  summarize(
    N = n(), 
    total_votes = sum(general_votes, na.rm = TRUE),
    d_votes = sum(ifelse(party == "DEM", general_votes, 0), na.rm = TRUE),
    r_votes = sum(ifelse(party == "REP", general_votes, 0), na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    other_votes = total_votes - d_votes - r_votes,
    r_prop = r_votes / total_votes,  
    winner = ifelse(r_votes > d_votes, "Republican", "Democrat")
  )
wi_results <- district_elections |>
  filter(state == "WI")
wi_results |>                  
  select(-state)
```

```{r, include = FALSE}
library(kableExtra)
```

Wisconsin's susceptibility to gerrymandering arises from its identity as a swing state characterized by contested elections. The state's distinct urban-rural disparity offers fertile ground for biased manipulation of electoral boundaries, prompting concerns regarding equal representation. To show these suspicions, we have leveraged data sourced from congressional district shape files for the 113th Congress, obtained from a UCLA website, and overlaid them onto the geographical borders of Wisconsin's congressional districts. The  table below provides a breakdown of the eight congressional districts alongside the respective percentages of Republican voters in each district.

```{r, echo=FALSE}
# distribution of total number of votes is narrow by design
# Proportion of Rep votes by district
wi_results |>
  select(district, r_prop, winner) |>
  arrange(desc(r_prop)) |>
  kable()
```

We observe a trend where Republican districts secure victories by narrow margins, contrasting with Democratic districts that either lack Republican voters entirely or win their districts uncontested. This suggests a potential instance of gerrymandering, where the state is strategically divided, concentrating Democratic voters into a single area to account for only one electoral representative, while Republicans dominate the majority of other districts.

These plots show the political party of each winning district as well as their shape and size relative to the state as well as the other districts. 

```{r, include =FALSE}
# Download congressional district shapefiles for the 113th Congress from a UCLA website (don't sweat the details too much)
src <- "http://cdmaps.polisci.ucla.edu/shp/districts113.zip"
lcl_zip <- fs::path(tempdir(), "districts113.zip")
download.file(src, destfile = lcl_zip)
lcl_districts <- fs::path(tempdir(), "districts113")
unzip(lcl_zip, exdir = lcl_districts)
dsn_districts <- fs::path(lcl_districts, "districtShapes")

# You can also downloaded zip file and uploaded it into R, but this uses a ton of space!
# dsn_districts <- fs::path("Data/districtShapes")

# read shapefiles into R as an sf object
st_layers(dsn_districts)

# be able to read as a data frame as well
districts <- st_read(dsn_districts, layer = "districts113") |>
  mutate(DISTRICT = parse_number(as.character(DISTRICT)))
head(districts, width = Inf)
class(districts)

# create basic plot with WI congressional districts
wi_shp <- districts |>
  filter(STATENAME == "Wisconsin")
wi_shp |>
  st_geometry() |>
  plot(col = gray.colors(nrow(wi_shp)))

# Append election results to geospatial data
wi_merged <- wi_shp |>
  st_transform(4326) |>
  inner_join(wi_results, by = c("DISTRICT" = "district"))
head(wi_merged, width = Inf)

# Color based on winning party
#   Note that geom_sf is part of ggplot2 package, while st_geometry is
#   part of sf package
```


```{r, echo = FALSE}
wi <- ggplot(data = wi_merged, aes(fill = winner)) +
  geom_sf(alpha = 0.5) +
  scale_fill_manual("Winner", values = c("blue", "red")) + 
  geom_sf_label(aes(label = DISTRICT), fill = "white") + 
  theme_void()
wi 
```


```{r, echo = FALSE}
library(leaflet)
pal <- colorNumeric(palette = "RdBu", domain = c(0, 1))

leaflet_wi <- leaflet(wi_merged) |>
  addTiles() |>
  addPolygons(
    weight = 1, fillOpacity = 0.7, 
    color = ~pal(1 - r_prop),   # so red association with Reps
    popup = ~paste("District", DISTRICT, "</br>", round(r_prop, 4))
  ) |>                          # popups show prop Republican
  setView(lng = -90, lat = 45, zoom = 6.5)
leaflet_wi
```


